### 2020-06-11 FDT2020/Introduction.tex
comments asnwered

\nocite{IPbook}
\section{Introduction}\label{chapter:intro}

\cindy{An original comment: There's material in this section (e.g. general discussion of combintatorial optimization) that's likely unnecessary for the journal paper. We probably want to start with what is now Section 2, but have some opening introduction giving the big picture.  Remember when we first mention rounding to be clear this is not numerical rounding. TODO: Add English statement for things currently only given in math. Add definitions of problems so reviewers and readers don't have to look them up.}
In combinatorial optimization the aim is to find the optimal solution in a discrete and
usually finite yet large set of solutions. For many specific combinatorial optimization problems such a solution can be found efficiently. For many others, finding optimal or in many cases near optimal solutions is NP-hard. A common approach to deal with such problems is relaxing the discrete solution set into a continuous set, where the optimization problem becomes tractable. Obtaining feasible solutions by means of such a relaxation requires an additional step of rounding the potentially fractional solution of the continuous relaxation into integer solutions.

\cindy{TODO: remove ``recall that'' for anything we haven't already defined in this paper.  Even though it's likely anyone reading this will care about IPs and integrality gaps, we should not assume they already know things we define the way we define them.}
In this paper, our focus is on linear relaxation of combinatorial
optimization problems. Combinatorial optimization was pioneered by Edmonds even before efficient algorithms for solving linear programming problems where introduced by Khachiyan \cite{Khachiyan} and later by Karmarkar \cite{Karmarkar1984}. For problems such as the \textsc{Minimum Cost Spanning Tree Problem} there are linear programming relaxations whose basic feasible solutions coincide with  integral solutions, i.e. spanning trees. For other problems the value of the linear programming relaxation provides a bound (lower bound for a minimization problem and upper bound for a maximization problem) on the optimal solution. A common and successful approach is to round these (potentially) fractional solutions into integer solutions for the combinatorial optimization problem at hand. The Integrality gap of a linear relaxation of an integer programming problem is the worst case ratio between the objective values of the discrete problem and the continuous problem. Equivalently, the integrality gap of the linear programming relaxation is a limit to the rounding approach: rounding a fractional solution into an integer solution incurs
a multiplicative cost proportional to the integrality gap. In this dissertation we study integrality gaps for different combinatorial optimization problems and introduce new rounding algorithms that imply bounds on their respective integrality gaps.

\section{Integrality Gap}
\iffalse{

@@

A polynomial time algorithm for proving an upper bound on integrality gap is called an LP-based approximation algorithm. For many well studied problems, we still do not know the exact integrality gap and the gap between the best known lower bound and the upper bound on the integrality gap are open. In some cases, there are known upper bounds, yet there is no known approximation algorithm, meaning that the proofs do not yield polynomial time algorithms.
}\fi

In this paper we focus on finding solutions to general Integer Linear Programs (IP). Integer Programming (and more generally Mixed Integer Linear Programming) can be used to model many practical optimization problems including scheduling, logistics and resource allocation. Recall that the set of feasible points for a pure IP (henceforth IP) is the set
\begin{equation}
S(A,b)= \{x\in \mathbb{Z}^{n}\;:\; Ax\geq b\}  \label{S}.
\end{equation}
If we drop the integrality constraints, we have the linear relaxation of set $S(A,b)$,
\begin{equation}
P(A,b) = \{x\in \mathbb{R}^{n}\;:\; Ax\geq b\}. \label{P}
\end{equation}

Let $I=(A,b)$ denote an instance. Then $S(I)$ and $P(I)$ denote $S(A,b)$ and $P(A,b)$, respectively. Given a linear objective function $c$, recall that an IP is $\min \;\{cx:\; x \in S(I)\}$. It is  NP-hard even to determine if an IP instance has a feasible solution~\cite{GareyJohnson}. However, intelligent branch-and-bound strategies allow commercial and open-source MILP solvers to give exact solutions (or near-optimal solutions with a provable bound) to many specific instances of NP-hard combinatorial optimization problems.

@@

Many researchers (see \cite{davids,vazirani}) have developed polynomial time LP-based approximation algorithms that find solutions for special classes of IPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. The approximation factor $C$ can be a constant or depend on the input parameters of the IP, e.g. $O(\log(n))$ where $n$ is the number of variables in the formulation of the IP (the dimension of the problem). However, for many combinatorial optimization problems there is a limit to such techniques based on LP relaxations, represented  by the {integrality gap} of the IP formulation. Recall that integrality gap $g(I)$ for instance $I$ is defined to be $g(I)= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}$. \cindy{Does the integrality gap always require positive objective coefficients?  Or is that an extra restriction in our definition?  Is there a reference that has that restriction?} For example consider the minimum cost 2-edge-connected multi-subgraph problem (2EC): Given a graph $G=(V,E)$ and $c\in \mathbb{R}^E_{\geq 0}$, 2EC asks for the minimum cost 2-edge-connected multi-subgraph of $G$. A linear programming relaxation for this problem known as the subtour elimination relaxation  is
 \begin{equation}\min \{cx: \sum_{e\in \delta(U)}x_e \geq 2 \mbox{ for } \emptyset \subsetneq U \subsetneq V,\; x\in [0,2]^{E}\}. 
 \end{equation}
 In this case the instance-specific integrality gap is the integrality gap of the subtour-elimination relaxation for the 2EC on graph with $n$ vertices. \cindy{We might need to define the problem or have a forward pointer.} \arash{See above}. Alexander et al. \cite{alexander2006integrality} showed the instance-specific integrality gap of the subtour elimination relaxation for the 2EC for instances of the problem with $n= 10$ is at most $\frac{7}{6}$.

@@

\begin{thm}[Carr, Vempala \cite{Carr2004}] \label{CV2}
	We have $g(I) \leq C$ if and only if for  $x\in P(I)$ there exists $\theta \in [0,1]^k$ where $\sum_{i=1}^{k}\theta_i =1$ and $\tilde{x}^i\in \dom(S(I))$ for $i\in [k]$ such that $\sum_{i=1}^{k}\theta_i \tilde{x}^i\leq Cx$.
\end{thm}
Recall that $\dom(P(I))$ is the set of points $x'$ such that there exists a point $x\in P$ with $x'\geq x$, also known as the dominant of $P(I)$. A polyhedron is of \textit{blocking type} if it is equal to its dominant. Theorem \ref{CV2} was first introduced by Goemans \cite{goemansblocking} for blocking type polyhedra. While there is an exact algorithm for problems with gap $1$ as stated in Proposition \ref{cara}, Theorem~\ref{CV2} is existential, with no construction.
%In contrast to Proposition \ref{cara} which implies exact algorithms for problems with a gap of 1, Theorem \ref{CV} does not imply an approximation algorithm, since it does not suggest how to find such a convex combination in polynomial time.
%This points to an interesting open question. 
% We show later, for $I$ with $g(I)<\infty$, the notation of dominant is in fact useful.
\iffalse

@@

In the \textsc{minimum vetex cover problem (VC)} we are given a graph $G=(V,E)$ and $c\in \mathbb{R}^E_{\geq 0}$. A subset of $U$ of $V$ is a \textit{vertex cover} if for $e\in E$ at least one endpoint of $e$ is in $U$. The goal in VC is to find the minimum cost vertex cover. The linear programming relaxation for VC is
\begin{equation}
\min \{cx \; : \; x_u + x_v \geq	 1 \text{ for } e=uv \in E, \; x\in [0,1]^{V}\}.
\end{equation}
The integrality gap of this formulation is exactly 2. \cindy{Move to active voice once we have an actual reference for this.} It is shown that it is UG-hard to approximte VC within any factor sctrictly better than 2 \cite{UGhardVC}. \cindy{We compare to small instances} We compare  FDT and the feasbility pump heuristic \cite{fp1} on the PACE 2019\footnote{ Parameterized Algorithms and Computational Experiments Challenge 2019: \url{https://pacechallenge.org/2019/}} challenge test cases \cite{PACE} and compare its performance with 
\subsubsection{Tree augmentation problem}
In the \textsc{Tree Augmentation Problem (TAP)} we are given a  graph $G=(V,E)$, a spanning tree $T$ of $G$. We also have a cost vector $c\in \mathbb{R}^{E\setminus T}_{\geq 0}$. A subset $F$ of $E\setminus T$ is called a \textit{feasible augmentation} if $(V,T\cup F)$ is a 2-edge-connected graph. In TAP we seek the minimum cost feasible augmentation. The natural linear programming relaxation for TAP is 
\begin{equation}
\min \{cx\; : \; \sum_{\ell \in \cov(e)} x_{\ell} \geq 1 \text{ for } e\in T, \; x\in [0,1]^{E\setminus T}\}.
\end{equation}
where $\cov(e)$ is set of edges $\ell \in E\setminus T$ such that $e$ is in the unique cycle of $T\cup \{\ell\}$. We call the LP above the cut-LP. The integrality gap of the cut-LP is known to be between $\frac{3}{2}$ \cite{32gaptap} and $2$ \cite{FJ81}. We create random fractional extreme points of the cut-LP and round them using FDT. For the instances that we create the blow-up factor is always below $\frac{3}{2}$ providing an upper bound for such instances.

>>>>

\nocite{IPbook}

\cindy{An original comment: There's material in this section (e.g. general discussion of combintatorial optimization) that's likely unnecessary for the journal paper. We probably want to start with what is now Section 2, but have some opening introduction giving the big picture.  Remember when we first mention rounding to be clear this is not numerical rounding. TODO: Add English statement for things currently only given in math. Add definitions of problems so reviewers and readers don't have to look them up.}
\iffalse{
\section{Introduction}\label{chapter:intro}

In combinatorial optimization the aim is to find the optimal solution in a discrete and
usually finite yet large set of solutions. For many specific combinatorial optimization problems such a solution can be found efficiently. For many others, finding optimal or in many cases near optimal solutions is NP-hard. A common approach to deal with such problems is relaxing the discrete solution set into a continuous set, where the optimization problem becomes tractable. Obtaining feasible solutions by means of such a relaxation requires an additional step of rounding the potentially fractional solution of the continuous relaxation into integer solutions.
}\fi
\cindy{TODO: remove ``recall that'' for anything we haven't already defined in this paper.  Even though it's likely anyone reading this will care about IPs and integrality gaps, we should not assume they already know things we define the way we define them.}

\section{Introduction}
In this paper we focus on finding solutions to general Integer Linear Programs (IP). Integer Programming (and more generally Mixed Integer Linear Programming) can be used to model many practical optimization problems including scheduling, logistics and resource allocation.

A common approach to Integer programming is considering their linear relaxation. For problems such as the \textsc{Minimum Cost Spanning Tree Problem} there are linear programming relaxations whose basic feasible solutions coincide with  integral solutions, i.e. spanning trees. For other problems the value of the linear programming relaxation provides a bound (lower bound for a minimization problem and upper bound for a maximization problem) on the optimal solution. A common and successful approach is to round these (potentially) fractional solutions into integer solutions for the optimization problem at hand. The Integrality gap of a linear relaxation of an integer programming problem is the worst case ratio between the objective values of the discrete problem and the continuous problem. Equivalently, the integrality gap of the linear programming relaxation is a limit to the rounding approach: rounding a fractional solution into an integer solution incurs
a multiplicative cost proportional to the integrality gap. In this dissertation we study integrality gaps for different combinatorial optimization problems and introduce new rounding algorithms that imply bounds on their respective integrality gaps.

\iffalse{

@@

A polynomial time algorithm for proving an upper bound on integrality gap is called an LP-based approximation algorithm. For many well studied problems, we still do not know the exact integrality gap and the gap between the best known lower bound and the upper bound on the integrality gap are open. In some cases, there are known upper bounds, yet there is no known approximation algorithm, meaning that the proofs do not yield polynomial time algorithms.
}\fi 
The set of feasible points for a pure IP (henceforth IP) is the set
\begin{equation}
S(A,b)= \{x\in \mathbb{Z}^{n}\;:\; Ax\geq b\}  \label{S}.
\end{equation}
If we drop the integrality constraints, we have the linear relaxation of set $S(A,b)$,
\begin{equation}
P(A,b) = \{x\in \mathbb{R}^{n}\;:\; Ax\geq b\}. \label{P}
\end{equation}
Let $I=(A,b)$ denote an instance. Then $S(I)$ and $P(I)$ denote $S(A,b)$ and $P(A,b)$, respectively. Given a linear objective function $c$, an IP is $\min \;\{cx:\; x \in S(I)\}$. It is  NP-hard even to determine if an IP instance has a feasible solution~\cite{GareyJohnson}. However, intelligent branch-and-bound strategies allow commercial and open-source MILP solvers to give exact solutions (or near-optimal solutions with a provable bound) to many specific instances of NP-hard combinatorial optimization problems.

@@

Many researchers (see \cite{davids,vazirani}) have developed polynomial time LP-based approximation algorithms that find solutions for special classes of IPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. The approximation factor $C$ can be a constant or depend on the input parameters of the IP, e.g. $O(\log(n))$ where $n$ is the number of variables in the formulation of the IP (the dimension of the problem). However, for many combinatorial optimization problems there is a limit to such techniques based on LP relaxations, represented  by the {integrality gap} of the IP formulation. The integrality gap $g(I)$ for instance $I$ is defined to be $g(I)= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}$. \cindy{Does the integrality gap always require positive objective coefficients?  Or is that an extra restriction in our definition?  Is there a reference that has that restriction?} For example consider the minimum cost 2-edge-connected multi-subgraph problem (2EC): Given a graph $G=(V,E)$ and $c\in \mathbb{R}^E_{\geq 0}$, 2EC asks for the minimum cost 2-edge-connected multi-subgraph of $G$. A linear programming relaxation for this problem known as the subtour elimination relaxation  is
 \begin{equation}\min \{cx: \sum_{e\in \delta(U)}x_e \geq 2 \mbox{ for } \emptyset \subsetneq U \subsetneq V,\; x\in [0,2]^{E}\}. 
 \end{equation}
 In this case the instance-specific integrality gap is the integrality gap of the subtour-elimination relaxation for the 2EC on graph with $n$ vertices. \cindy{We might need to define the problem or have a forward pointer.} \arash{See above}. Alexander et al. \cite{alexander2006integrality} showed the instance-specific integrality gap of the subtour elimination relaxation for the 2EC for instances of the problem with $n= 10$ is at most $\frac{7}{6}$.

@@

\begin{thm}[Carr, Vempala \cite{Carr2004}] \label{CV2}
	We have $g(I) \leq C$ if and only if for  $x\in P(I)$ there exists $\theta \in [0,1]^k$ where $\sum_{i=1}^{k}\theta_i =1$ and $\tilde{x}^i\in \dom(S(I))$ for $i\in [k]$ such that $\sum_{i=1}^{k}\theta_i \tilde{x}^i\leq Cx$.
\end{thm}
We denote by $\dom(P(I))$ the set of points $x'$ such that there exists a point $x\in P$ with $x'\geq x$, also known as the dominant of $P(I)$. A polyhedron is of \textit{blocking type} if it is equal to its dominant. Theorem \ref{CV2} was first introduced by Goemans \cite{goemansblocking} for blocking type polyhedra. While there is an exact algorithm for problems with gap $1$ as stated in Proposition \ref{cara}, Theorem~\ref{CV2} is existential, with no construction.
%In contrast to Proposition \ref{cara} which implies exact algorithms for problems with a gap of 1, Theorem \ref{CV} does not imply an approximation algorithm, since it does not suggest how to find such a convex combination in polynomial time.
%This points to an interesting open question. 
% We show later, for $I$ with $g(I)<\infty$, the notation of dominant is in fact useful.
\iffalse

@@

In the \textsc{minimum vetex cover problem (VC)} we are given a graph $G=(V,E)$ and $c\in \mathbb{R}^E_{\geq 0}$. A subset of $U$ of $V$ is a \textit{vertex cover} if for $e\in E$ at least one endpoint of $e$ is in $U$. The goal in VC is to find the minimum cost vertex cover. The linear programming relaxation for VC is
\begin{equation}
\min \{cx \; : \; x_u + x_v \geq	 1 \text{ for } e=uv \in E, \; x\in [0,1]^{V}\}.
\end{equation}
The integrality gap of this formulation is exactly 2 \cite{davids}. \cindy{Move to active voice once we have an actual reference for this.} \arash{I guess we can site a general approx book,} It is shown that it is UG-hard to approximte VC within any factor sctrictly better than 2 \cite{UGhardVC}. \cindy{We compare to small instances} We compare  FDT and the feasbility pump heuristic \cite{fp1} on the the small instances of PACE 2019\footnote{ Parameterized Algorithms and Computational Experiments Challenge 2019: \url{https://pacechallenge.org/2019/}} challenge test cases \cite{PACE}. 
\subsubsection{Tree augmentation problem}
In the \textsc{Tree Augmentation Problem (TAP)} we are given a  graph $G=(V,E)$, a spanning tree $T$ of $G$. We also have a cost vector $c\in \mathbb{R}^{E\setminus T}_{\geq 0}$. A subset $F$ of $E\setminus T$ is called a \textit{feasible augmentation} if $(V,T\cup F)$ is a 2-edge-connected graph. In TAP we seek the minimum cost feasible augmentation. The natural linear programming relaxation for TAP is 
\begin{equation}
\min \{cx\; : \; \sum_{\ell \in \cov(e)} x_{\ell} \geq 1 \text{ for } e\in T, \; x\in [0,1]^{E\setminus T}\}.
\end{equation}
where $\cov(e)$ is set of edges $\ell \in E\setminus T$ such that $e$ is in the unique cycle of $T\cup \{\ell\}$. We call the LP above the cut-LP. The integrality gap of the cut-LP is known to be between $\frac{3}{2}$ \cite{32gaptap} and $2$ \cite{FJ81}. We create random fractional extreme points of the cut-LP and round them using FDT. For the instances that we create the blow-up factor is always below $\frac{3}{2}$ providing an upper bound for such instances.
